\section{Discussion}

Our results show that the point system for a game of Yahtzee is not necessarily
based on the expected values for each of the possible scoring criteria.
For example, while the expected value of a Yahtzee is about 2.3, the expected
value for a full house (which the game rules give less points to than rolling
a Yahtzee) is 7.8, slightly more than triple the expected value for a Yahtzee.
This shows that while the game's scoring system encourages a certain style of
play (namely, trying for Yahtzees), a much higher return can be expected from
rolling for different combinations (such as a full house).

Now from the results the game follows an inconsistent point to probability curve.
The expected value for one to get from a Yahtzee, the highest scoring roll,
ends up with only a 2.3 point average while one going for a number like 3 has an
expected value of 6.3 points on average. So, striving to roll a Yahtzee in the
game is actually one of the poorer strategies to win the game.  Therefore, this
is not a true game a chance and some rolls have a greater value (such as with
just rolling for a number).

Rolling for twos when compared to rolling for sixes has the same probability,
but rolling for sixes has a greater point return than rolling for twos.  The
game is not built to be purely chance for the players, and as such there are
some real skills and decision making to gain victory in the game of Yahtzee.
The game provides experienced players to gain an edge to a seemingly luck based
game of rolling some dice.

However, even an experienced player can lose to an inexperienced player because
of the element of chance in the dice which can still result in a game swinging
in favor of the inexperienced player every once and awhile.  So if we compare
a person who rolls to get yahtzees to a person who rolls for full houses then we
see an average of 5.547 points of difference between the two players per round
and 72.0421 points per game which is a huge disparity when we look at only
rolling for a different type of result.
